---
title: "NLP"
output: html_document
---

The problem proposed in this notebook is to analyze a subset of tweets from the premier league and analyze it using sentiment analysis to determine whether the impact of social networks influence on the teams and player results.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# This R environment comes with many helpful analytics packages installed
# It is defined by the kaggle/rstats Docker image: https://github.com/kaggle/docker-rstats
# For example, here's a helpful package to load

library(tidyverse) # metapackage of all tidyverse packages

# Input data files are available in the read-only "../input/" directory
# For example, running this (by clicking run or pressing Shift+Enter) will list all files under the input directory

list.files(path = "../input")

# You can write up to 20GB to the current directory (/kaggle/working/) that gets preserved as output when you create a version using "Save & Run All" 
# You can also write temporary files to /kaggle/temp/, but they won't be saved outside of the current session

```

We set a seed to get the same results in different executions

```{r}
set.seed(123)  
```

## Read data

First of all we will need to download the csv files from this link <https://www.kaggle.com/wjia26/epl-teams-twitter-sentiment-dataset> and unzip the file in the data folder.

We will read the csv files, we will use this dataset <https://www.kaggle.com/wjia26/epl-teams-twitter-sentiment-dataset>. It contains tweets of all teams of the premier from 09-07-2020 to 20-09-2020.

```{r}
data <- read.csv('data/2020-07-09 till 2020-09-19.csv')
#tweets2 <- read.csv('data/2020-09-20 till 2020-10-13.csv') 
#data <- rbind(tweets1, tweets2)
head(data)

```

As we can see we have some irrelevant information like:

-   group_name

-   screenname

-   twitter_id

-   username

We will pop this info from the dataframe

```{r}
data$group_name <- NULL
data$screenname <- NULL
data$twitter_id <- NULL
data$username <- NULL

head(data)

```

Now we will proceed to clean the text.

## Clean text

First of all we check if the encoding is correct

```{r}
#install.packages('utf8')
library(utf8)
linesQ <- data$text
#Chack encoding
linesQ[!utf8_valid(linesQ)] #character(0) ==> All lines are made of correct UTF-8 characters
```

We get character(0) that means all lines are mead of correct UTF-8 characters. We can read every character so now we will start cleaning our text.

We will start transforming emojis into text. This takes a while.

```{r}
library(textclean)
linesQ <- replace_emoji(linesQ)
linesQ <- replace_emoticon(linesQ)
print(linesQ[1:5])
```

We will remove from the text unnecessary info like URL, mentions, hashtgas...

```{r}
#Remove RT
linesQ <- gsub("(RT|via)((?:\\b\\W*@\\w+)+)", "", linesQ)
#Remove mentions
linesQ <- gsub("@\\w+", "", linesQ)
#Remove links
linesQ <-  gsub(" ?(f|ht)tp(s?)://(.*)[.][a-z]+", "", linesQ)
#Remove non alphanumeric chars
linesQ <- gsub("[^a-zA-Z0-9 ]", "", linesQ)
#Remove punctuation
linesQ <- gsub("amp ", "", linesQ)
#Remove Na
linesQ <- linesQ[!is.na(linesQ)]
#Spaces and tabs cleaning
linesQ <- gsub("[ \t]{2,}", "", linesQ)
linesQ <- gsub("^\\s+|\\s+$", "", linesQ)
#Convert to lower case
linesQ <- tolower(linesQ)

head(linesQ, 5)
```

We check everything seems to be alright.

## Data analysis using library tm

First we create a corpus and apply some cleaning too

```{r}
library(tm)
linesQCorpus <- Corpus(VectorSource(linesQ))
corpus_clean <- tm_map(linesQCorpus, tolower)
corpus_clean <- tm_map(corpus_clean, removeNumbers)
corpus_clean <- tm_map(corpus_clean, removePunctuation)
corpus_clean <- tm_map(corpus_clean, stripWhitespace)
#Remove stopwords
corpus_clean <- tm_map(corpus_clean, removeWords, stopwords('english'))
```

We create the document term matrix

```{r}
dtm <- DocumentTermMatrix(corpus_clean, control = list(minWordLength = 1, stopwords = TRUE))
inspect(dtm)
```

And the most frequent terms are:

```{r}
head(findFreqTerms(dtm, lowfreq=10), 40)
```

This will be later visualized in a wordcloud

## More EDA using spacyR and other resources

This takes a long time.

```{r}
library(spacyr)
#spacy_install()
spacy_initialize(model = "en_core_web_sm")
#Gets sentences from paragraphs
phrases <- spacy_tokenize(linesQ, #If you use quanteda you can use
# corpus_reshape(corpus, to = "sentences"))
#Taks a while.
what="sentence" #By default remove_separators = TRUE
# (removes trailing spaces)
)
v_phrases <- unlist(phrases)
numphrases <- length(v_phrases) #8,975 sentences
sum(v_phrases=="") #1

```

What about the length of the sentences?

```{r}
#A simple histogram will do fine
hist(nchar(v_phrases),
main = "Histogram of sentence size",
xlab = "Sentece size (number of characters)",
ylab = "Ocurrences",
xlim=c(0,1000)
)
```

As it was expected there are a lot of sentences with very few characters, one of the reasons is the Twitter limit of characters per tweet.

And what about the most frequent terms?

```{r}
  tokens <- spacy_tokenize(linesQ
  #Parameters asigned by default:
  #remove_punct = FALSE, punt symbols are tokens
  #remove_url = FALSE, url elements are tokens
  #remove_numbers = FALSE, numbers are tokens
  #remove_separators = TRUE, spaces are NOT tokens
  #remove_symbols = FALSE, symbols (like €) are tokens
  )#Returns a list
  v_tokens <- unlist(tokens)
  v_tokens[1:10]

```

```{r}
#As a simple plot
plot(head(sort(table(v_tokens), decreasing = TRUE), n = 10),
xlab = "Token",
ylab = "Ocurrences"
)

```

Most of these tokens are stopwords.

We can see symbols \< \> are the most used these are related to emojis

```{r}
#spacy_finalize() #Do not forget this
```

```{r}
data %>%
  # UCT time in hh:mm format
  mutate(created_at=substr(created_at, 12, 16)) %>%
  count(created_at) %>%
  ggplot(aes(x=as.numeric(as.factor(created_at)), y=n, group=1)) +
  geom_line(size=1, show.legend=FALSE) +
  geom_vline(xintercept=7, colour="red") +
  labs(x="UCT time (hh:mm)", y="Number of Tweets") + 
  theme_bw() +
  scale_x_continuous(breaks=c(1,301,601,901,1201,1440),
                     labels=c("00:00","05:00","10:00","15:00","20:00","23:59") )

```

Most of the tweets are tweeted during gametime.

Number of words per tweet.

```{r}
data %>%
  mutate(words_per_tweet=sapply(strsplit(text, " "), length)) %>%
  ggplot(aes(x=words_per_tweet)) +
  geom_histogram(bins=10, show.legend=FALSE) +
  xlim(c(0,40)) +
  theme_bw() +
  labs(x="Words", y="Frequency")
```

## Sentiment analysis

We will start the sentiment analysis part visualizing what are the most frequent positive and negative words. We will use bing lexicon(<https://www.cs.uic.edu/~liub/FBS/sentiment-analysis.html>).

```{r}
library(tidytext)
library(wordcloud)
library(wordcloud2)
library(reshape2)

# Tokens
tokens <- data %>%  
  unnest_tokens(word, text) %>%
  select(word)

# Positive and negative words 
tokens %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort=TRUE) %>%
  acast(word ~ sentiment, value.var="n", fill=0) %>%
  comparison.cloud(colors=c("#F8766D", "#00BFC4"), max.words=200)
```

Using nrc lexicon(<http://saifmohammad.com/WebPages/NRC-Emotion-Lexicon.htm>) we can see the frequency of words according to the sentiment expressed.

```{r}
# Sentiments and frequency associated with each word  
sentiments <- tokens %>% 
  inner_join(get_sentiments("nrc")) %>%
  count(word, sentiment, sort=TRUE) 

# Frequency of each sentiment
ggplot(data=sentiments, aes(x=reorder(sentiment, n, sum), y=n)) + 
geom_bar(stat="identity", aes(fill=sentiment), show.legend=FALSE) +
labs(x="Sentiment", y="Frequency") +
theme_bw() +
coord_flip()

```

And using AFINN lexicon (<http://www2.imm.dtu.dk/pubdb/pubs/6010-full.html>) we will get the words with highest and lowest sentiment scores.

```{r}
library(gridExtra)
# Positive and negative words 
top_positive <- tokens %>% 
  inner_join(get_sentiments("afinn")) %>%
  count(word, value, sort=TRUE) %>%
  arrange(desc(value)) %>%
  head(n=10) %>%
  ggplot(aes(x=reorder(word, value), y=value)) +
  geom_bar(stat="identity", fill="#00BFC4", colour="black") +
  theme_bw() +
  labs(x="Positive words", y="value") +
  coord_flip() 

top_negative <- tokens %>% 
  inner_join(get_sentiments("afinn")) %>%
  count(word, value, sort=TRUE) %>%
  arrange(value) %>%
  head(n=10) %>%
  ggplot(aes(x=reorder(word, -value), y=value)) +
  geom_bar(stat="identity", fill="#F8766D", colour="black") +
  theme_bw() +
  labs(x="Negative words", y="value") +
  coord_flip() 

grid.arrange(top_positive, top_negative,
             layout_matrix=cbind(1,2))
```

As we can see the negative words are very rude getting a very high negative value.

We can take advantage of the dataset and use the polarity field that already express a sentiment score of each tweet and the file_name variable that associates each tweet to a team to see which are the teams who receive the best score.

```{r}
teams_score <-aggregate(data$polarity, by=list(File=data$file_name), FUN=sum)
ordered <- teams_score[order(teams_score$x),]
#As a simple plot
p<-ggplot(data=ordered, aes(x=File, y=x)) +
  geom_bar(stat="identity", fill="steelblue")+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90))
p
```

## Named Entity Recognition

Another interesting question to solve is. Is there a team or a player that receives more mentions than anothers? We will apply NER to solve this question. This takes a lot of time.

```{r}
library(spacyr)
#spacy_install()
spacy_initialize(model = "en_core_web_sm")
entities <- spacy_parse(linesQ, lemma = FALSE, entity = TRUE) %>%
    entity_extract() %>%
    group_by(doc_id) %>%
    summarize(ner_words = paste(entity, collapse = ", "))
spacy_finalize() #Do not forget this
```

Now let's count the number of ocurreces.

```{r}
solution<-as.data.frame(entities)
plot(
  
  head(sort(table(unlist(solution)), decreasing = TRUE), n = 10),
xlab = "Entities",
ylab = "Ocurrences"
)
```
